---
title: "Data Science Assignment 02"
author: "Zak Zerhouni"
date: today
format:
  html:
    embed-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

## Instructions

Complete the following exercises **using only the tidyverse packages** (e.g., `dplyr`, `tidyr`, `ggplot2`, `readr`, `stringr`, `forcats`, `lubridate`, `purrr`, etc.)

**Important:** This assignment is intended to assess your proficiency with the **tidyverse**. Solutions that use **base R** or **packages outside the tidyverse** to solve the problems will **not receive credit** unless explicitly permitted in the question.

1.  Load the necessary **tidyverse** libraries.
2.  Answer each question in a separate R code chunk.
3.  Provide a detailed explanation of your approach for each solution.
4.  Render your R Markdown document to an HTML file.
5.  Submit the rendered HTML file to Canvas.

# Questions:

## Task 1: Read data using a relative path

We first load the necessary `tidyverse` package. Then we user `read_csv` from `readr` to store the data from the `collegedata.csv` file using a relative path. In order to form the relative path string a period, `.`, is used to point to the current directory. 

```{r}
#| warning: false
require(tidyverse)

suppressMessages(
  college.data <- readr::read_csv(file = "./data/collegedata.csv")
)
head(college.data)
```

## Task 2: Determine rows and columns

The number of rows and columns in a data set can be displayed using the `glimpse` function in the `dplyr` package. It also displays information about the data type of each column.

```{r}
dplyr::glimpse(college.data)
```

## Task 3: Use tidyverse to select INSTNM, SAT_AVG, and ADM_RATE

The `select` function of `dplyr` is used to save a subset of the data only including the columns INSTNM, SAT_AVG, and ADM_RATE.

```{r}
college.subset <- college.data |> select(INSTNM,SAT_AVG,ADM_RATE)
head(college.subset)
```


- Task 4: Create a scatterplot for SAT_AVG and ADM_RATE (Remove NA)

## Submission

- After completing **each question**, save your Quarto file, render it, then **stage and commit** your changes.

- Each commit should:

  - Represent **one logical change**.
  - Have a **short, descriptive** commit message.
  - Be written in the **present tense** (imperative mood).

  **Examples:**

  **Bad:** `Fix bug in the project` *Too vague—it does not explain which bug was fixed.*

  **Bad:** `Fixed crash on login screen when password is empty` *Uses the past tense.*

  **Good:** `Fix crash on login screen when password is empty`

- Include well-commented code and clear explanations for each solution.

- Make **at least one meaningful commit per question/part** (a minimum of **4 meaningful commits**).

- **Push your work to your** GitHub\*\*. ane post the link on the text bar of the Canvas assignment.

- Assignments that are **not pushed to GitHub** will receive **no more than 60% of the available points**.

- Submit the rendered HTML file to **Canvas**. Assignments that are **not submitted to Canvas** will also receive **no more than 60% of the available points**.

- If you forget to submit your work to **GitHub** or **Canvas**, you have **up to three (3) days** after the due date to correct the submission. A **10% late penalty** will be applied to the assignment. **After three days, no submissions or corrections will be accepted.**

- **No corrections, re-submissions, or missing submissions will be accepted after the assignment has been graded**, including forgotten GitHub or Canvas submissions.

**Good luck!**
